---
title: "Homework 5"
subtitle: "ECON 470, Spring 2025"
author: "Molly Catlin"
format:
  pdf:
    output-file: "catlin_molly_hwk5_s2"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
    df-print: kable
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, stringr, readxl, 
               data.table, gdata, scales, kableExtra, modelsummary, fixest)
```

Here is a link to my repository: {https://github.com/mollyjc02/Homework_5.git}


\newpage 


```{r}
#| include: false
#| eval: true

library(here)
load(here("submission_2", "results", "hwk5_workspace.RData"))
```


\newpage 


## 1. Plot the share of the adult population with direct purchase health insurance over time. 
 

```{r} 
#| echo: false
#| label: fig-dir-share
#| fig-cap: "Share of Individuals with Direct Purchase" 

direct.share.plt
```


\newpage 


## 2. Discuss the reduction in direct purchase health insurance in later years. Can you list a couple of policies that might have affected the success of the direct purchase insurance market? 
answer here 


\newpage 


# 3. Plot the share of the adult population with Medicaid over time. 


```{r} 
#| echo: false
#| label: fig-med-share
#| fig-cap: "Share of Individuals with Medicaid"

medicaid.share.plt
```


\newpage 


# 4. Plot the share of uninsured over time, separately by states that expanded Medicaid in 2014 versus those that did not. Drop all states that expanded after 2014. 


```{r} 
#| echo: false
#| label: fig-unisured 
#| fig-cap: "Average Uninsured by Medicaid Expansion"

uninsured.plot 
```


\newpage 


# 5. Calculate the average percent of uninsured individuals in 2012 and 2015, separately for expansion and non-expansion states. Present your results in a basic 2x2 DD table. 


```{r} 
#| echo: false
#| label: tab-dd
#| tbl-cap: "DD Table for Medicaid Expansion"

knitr::kable(dd.table, digits = 3, caption = "Difference-in-Differences Table of Uninsured Rates")
```


\newpage


# 6. Estimate the effect of Medicaid expansion on the uninsurance rate using a standard DD regression estimator, again focusing only on states that expanded in 2014 versus those that never expanded.


```{r}
#| echo: false
#| label: tab-dd-reg
#| tbl-cap: "DD Estimates for Medicaid Expansion"

modelsummary(dd.est,
             coef_rename=c("postTRUE" = "Post 2014","expand_everTRUE"="Expand",
                           "treat" = "Post x Expand"),
             gof_omit='DF|F|Lik|AIC|BIC|Adj')
```
